A geo-referenced micro-data set of real estate listings for Spain’s three largest cities

### 2022-02-18 EPB/EPB.Rmd
Version EPB con inglés corregido

---
title: A geo-referenced micro-data set of real estate listings for the three largest Spanish cities
runninghead: Rey-Blanco \emph{et al}.
author:
- name: D. Rey-Blanco
  num: 1
- name: P. González-Arbues
  num: 1
- name: F. López
  num: 2
- name: A. Páez*
  num: 4
address:
- num: 1
  org: Idealista, Plaza de las Cortes 5, 28014 Madrid, Spain
- num: 2
  org: Facultad de CC de la Empresa, C/ Real, 3. 30201 Cartagena, Murcia, Spain
- num: 3
  org: School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada
corrauth: "Antonio Páez, School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada."
email: paezha@mcmaster.ca
abstract: "This data article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain. These data were originally published on idealista.com real estate website. The observations are obtained for the three largest Spanish cities: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations) and Valencia (n = 33,622 observations). The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics. The listings were enriched with official information (building year of construction and built quality materials grade) plus other relevant geographical features such as distance to urban points of interest. Along with real estate listings, the data product also includes neighborhood boundaries for each city. The data product is offered in the form of a fully documented R package. This open data product is available for scientific and educational purposes, in particular for geo-spatial studies."
keywords: Housing market; idealista.com; geo-referenced data; open data; hedonic price analysis; Spain
classoption:
  - Royal
  - times
bibliography: bibEPB.bib
bibliographystyle: sageh

@@

El interés pos el desarrollo de modelos hedónicos de preicción del precio de lvivienda que incluyen la compenente espaciál,  y en general de la importancia de la geografía para realizar un correcto analisis del mercado inmobiliario, ha siod un tópico de creciente interes [@lopez2015; @crespo2013local].

Por otra parte la posibilidad de disponer de datos a nivel micro/urbano de la vivienda ha sido dificil y en algunos casso se ha tenido que recurrir a dudosos procesos de webscraping para poder obtener grandes volúmnees de informacion que permitan. En muchos casos estos procesos de webscrappin pueden contener datos faltantes, errores, datos duplicados etc,

El creciente interés por datos en abierto [@arribas2021]

The main objective of this paper is present a sort description of the an open data set of a geo-referenced dwelling listings.

This data set is excelente para aplicar modelos hedónicos con efectos espaciales, identificar submercados de vivienda, aplicar técnicas de machine learning.

La importancia de disponer de datos en abierto

Pocos datos en abierto de precios de vivienda

Pocos con informacion georeferenciada a nivel de punto

The data set is distributed in the form of an `R` package, named **idealista18** available from https://github.com/paezha/idealista18 . All spatial objects such as polygons and points are distributed as simple features objects (class `sf` in R). The data has been provided by Idealista, the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal.

POR QUE HACEN UNA CONTRIBUCIÓN A LA COMUNIDAD URBAN ANALITICS

@@

The open data set 'idealista18' is composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid and Valencia. For each city, dwelling listings, neighborhood polygons and a set of points of interest (POI) has been included in the R package. The next subsections describe each data set.

@@

The dwelling listing of each city, include a set of characteristics of each dwelling that was published on idealista real state website (https://www.idealista.com/) and are included in 'idealista18' package as an sf object [@Pebesma]. The data set corresponding to each city is named with the name of the city followed by '_Sale'. Each of these files contains the complete set of listings, corresponding to the four quarters of year 2018. The record counts for each city in 2018 are: 94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for Valencia. Table \ref{tab:number-ads} show the number of ads included in the data set for city and quarter. Note that is possible that the same dwelling can be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter. The variable ASSETID, included in each sf object is the unique identifier of the dwelling.

```{r number-ads, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
H <- t(cbind(table(Barcelona_Sale$PERIOD),table(Madrid_Sale$PERIOD),table(Valencia_Sale$PERIOD)))
H <- cbind(H,rowSums(H))
text_tbl <- data.frame(
City = c("Barcelona","Madrid","Valencia"),
H,
check.names = FALSE)
colnames(text_tbl) <- c("City", "First", "Second ", "Thirdr", "Fourth","Total ads")
xtable::xtable(text_tbl, digits=0, caption = "Number of dwelling  listing ads for each city and quarter. \\label{tab:number-ads}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F) %>%
column_spec(1, width = "4em", italic = FALSE) %>%
column_spec(2, width = "3em") %>%
column_spec(3, width = "3em", latex_column_spec = "c") %>%
column_spec(4, width = "3em", latex_column_spec = "c") %>%
column_spec(5, width = "3em", latex_column_spec = "c") %>%
column_spec(6, width = "2em", latex_column_spec = "c")
```

Each record of the dwelling listing contains a set of indoor characteristics supplied by advertiser (e.g. price, surface, rooms, basic features, etc) beside with the exact localization of the dwelling (see Section \ref{sec:anonymizing}). Table \ref{tab:variables} list some of the main indoor variables included in the dwelling listing with a short description and the mean value of each variable. This dwelling listing was enriched with a number of additional attributes from the Spanish cadastre [@Catastro]. Cadastral information is described in Table \ref{tab:variables}, including the the prefix CAD in the variable name. Cadastral features assignment is done by assigning the features of the nearest parcel to the coordinates. The year of construction of dwelling was revised, given that original year of construction from listings are entered by users in the web site, therefore subject to errors and incomplete (a 40\% missing rate). To remove any issue we assign cadastral construction year from the nearest cadastral parcel whenever value has an outstanding value (date is after publication date or year of construction is before 1500) or when the field value was missing. Additionally, the distance of each dwelling to three urban points of interest was included in the data set: distance to city center, distance to the closed metro station and distance to the main street (Diagonal street for Barcelona, Castellana street for Madrid and Blasco Ibañez street for Valencia). The last rows of Table \ref{tab:variables} show the mean values of this variables.

```{r table-variables, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
dummyvariables <- c(3,4,5,6,7,27,30,31,32,33,37,38,39)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("ksjdhfal","Asking price per m^2 (euros)","Surface (m^2)",
                       "Number of bedrooms","Number of bathrooms",
                       "Construction year (advertiser)","Construction year (cadastre)",
                       "Max build floor","Dwelling count in the building",
                       "Cadastral quality. 0 Best-10 Worst",
                       "Distance to city center","Distance to subway station","Distance to main street"),
# Barcelona
Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables]), na.rm = TRUE),
# Madrid
Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables]), na.rm = TRUE),
# Valencia
"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables]), na.rm = TRUE),
check.names = FALSE)

xtable::xtable(text_tbl, caption = "List of quantitative variables included in the dwelling listing for the three Spanish cities. See the help facility in the \\textbf{idealista18} R package for details and formal definitions. Some variables has been excluded of this table for save space, check the full list in  \\textbf{idealista18} R package. \\label{tab:variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F, font_size=8) %>%
column_spec(1, width = "13em", italic = FALSE) %>%
column_spec(2, width = "14em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```

To conclude the description of the dwelling listing, Table \ref{tab:Dummy-variables} include the more relevant variables that include information about basic characteristics of the dwellings. All variables listed in this table are dummy variables and the percentaje of dwelling with the characteristis

```{r table-dummy-variables, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
dummyvariables <- c(8,9,10,12,15,16,17,18,19,20,21,22,23,24,25,26,34,35,36)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("=1 if has terrace","=1 if has lift",
                       "=1 if has air conditioning","=1 if has parking",
                       "=1 if has north orientation","=1 if has south orientation",
                       "=1 if has east orientation","=1 if has west orientation",
                       "=1 if has boxroom","=1 if has wardrobe",
                       "=1 if has swinningpool","=1 if has doorman",
                       "=1 if has garden","=1 if is duplex",
                       "=1 if is studio","=1 is in the top floor",
                       "=1 if is new","=1 is second hand to be restored","=1 is second hand good conditions"),
# Barcelona
Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
# Madrid
Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables])),
# Valencia
"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables])),
check.names = FALSE)

xtable::xtable(text_tbl, caption = "List of dummy variables with the percentage of dwelling with a specific characteristic. See the help facility in the \\textbf{idealista18} R package for details and formal definitions. Some dummy variables has been excluded of this table for save space. \\label{tab:Dummy-variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F, font_size = 8) %>%
column_spec(1, width = "12em", italic = FALSE) %>%
column_spec(2, width = "14em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```

@@

The second block of data included in the 'idealista18' R package are the spatial features of the three cities divided in neighborhoods. Figure \ref{fig:all-polygons} shows the different neighborhoods for the three cities. The boundaries are based on the official boundaries but slightly adapted by idealista\footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}. In practical terms we can assume they are the same, since the website simply collapses areas when they are sufficiently small in terms of number of ads. In the case of Madrid they just collapse four areas into two new ones.

A total of

```{r, echo = FALSE, fig.width=5, fig.cap="\\label{fig:all-polygons} (A) Weekly intra-province"}
library(ggplot2)
library(gridExtra)
plot1 <- ggplot(data=Barcelona_Polygons) + geom_sf(color = "black",size=.2) + 
  theme_bw(base_size = 2) + 
  ggtitle("                Barcelona") +
  theme(plot.title = element_text(size = 10),
        panel.border = element_blank(),
    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
plot2 <- ggplot(data=Madrid_Polygons) + geom_sf(color = "black",size=.2) + 
  theme_bw(base_size = 2) +
   ggtitle("                Madrid") +
    theme(plot.title = element_text(size = 10),panel.border = element_blank(),
    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
plot3 <- ggplot(data=Valencia_Polygons) + geom_sf(color = "black",size=.2) + 
  theme_bw(base_size = 2) +
   ggtitle("                Valencia") +
    theme(plot.title = element_text(size = 10),panel.border = element_blank(),
    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
grid.arrange(plot1, plot2, plot3, ncol = 3, nrow = 1)
```

There are a total of 73 neigborhoods in Barcelona, 135 in Madrid and 73 in Valencia. The sf object include an unique identifier (LOCATIONID) and the neighborhood name (LOCATIONNAME).

@@

The last block of data included in the data package is a set of Point of Interest of each city in `R` list R objetc. These lists include three elements: (i) the coordinates of the city center (see that identify the central business district); (ii) a set of points that define the main street of each city; and (iii) the coordinates of metro stations.

# Anonymizing the data set {#sec:anonymizing}

To comply with Spanish regulations, two variables are slightly modified to preserve their anonymity. A masking process is apply to asking prices and localization.

Whit respect to the asking prices, the original values are obfuscated with the addition or subtraction of a random percentage of their original values ranging from -2.5\% to +2.5\%. Since asking prices are usually multiples of 1000, after the first price modification, the prices was aligned to multiples of 1000.

With respect to the dwelling localization, a spatial masking process was implemented with the intention of keeping spatial properties of the original data set. The coordinates of each listing were displaced using a stochastic procedure. Effectively, the listings were recorded using coordinates contained in a maximum and minimum displacement circles, as shown in Figure \ref{fig:points-moved-image}. To preserve membership in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates are in the original neighborhood of the listing.

```{=tex}
\begin{figure}[!ht]
  \caption{Masking coordinates. Spatial range}
  \centering
  \includegraphics[width=6cm, height=4.9cm]{EPB_files/points-moved-image.png}
  \label{fig:points-moved-image}
\end{figure}
```

The algorithm \ref{algo:coordinates-displacement} iteratively displaces the coordinates of each listing with a minimum distance and a maximum distance with the restriction that the new coordinates do not fall in a different neighborhood. This ensures that neighborhood attributes are preserved.

@@

Figure \ref{fig:coordinates-displacement} shows the histogram of displacements in meters for all listings in the city of Valencia; the average distance between the original and masked coordinates is 45 meters.

```{=tex}
\begin{figure}[!ht]
  \caption{Masking coordinates. Spatial range}
  \centering
  \includegraphics[width=6cm, height=4.9cm]{EPB_files/coordinates-valencia.png}
  \label{fig:points-moved-image}
\end{figure}
```

## The geo-referenced micro-data set

Spatial objects include geodetic coordinates using the \emph{EPSG:4326} coordinate reference system.

>>>>

---
title: A geo-referenced micro-data set of real estate listings for Spain’s three largest cities
runninghead: Rey-Blanco \emph{et al}.
author:
- name: D. Rey-Blanco
  num: 1
- name: P. González-Arbues
  num: 1
- name: F. López
  num: 2
- name: A. Páez*
  num: 3
address:
- num: 1
  org: Idealista, Plaza de las Cortes 5, 28014 Madrid, Spain
- num: 2
  org: Facultad de CC de la Empresa, C/ Real, 3. 30201 Cartagena, Murcia, Spain
- num: 3
  org: School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada
corrauth: "Antonio Páez, School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada."
email: paezha@mcmaster.ca
abstract: "This article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain. These data were originally published on the idealista.com real estate website. The observations were obtained for the three largest cities in Spain: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations), and Valencia (n = 33,622 observations). The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics. The listings were enriched with official information from the Spanish cadastre (e.g., building material quality) plus other relevant geographical features, such as distance to urban points of interest. Along with the real estate listings, the data product also includes neighborhood boundaries for each city. The data product is offered as a fully documented R package and is available for scientific and educational purposes, particularly for geo-spatial studies"
keywords: Housing market; idealista.com; geo-referenced data; point-level data; open data; Spain
classoption:
  - Royal
  - times
bibliography: bibEPB.bib
bibliographystyle: sageh

@@

Interest in the characteristics of the housing market and housing prices has been a growing area of research in recent decades, generating a vast amount of theoretical and empirical literature. Including the spatial component to analyze the real estate market and incorporating geographic variables has significantly improved the understanding of this market. But to really understand the characteristics of the housing market, it is essential to have information/data at the point level. Therefore, it is becoming common for spatial analysis of urban environments to be developed with geo- referenced micro-data sets [@lopez2015]. However, the availability of this type of open data at the point level is limited, and not many data sets contain latitude/longitude coordinates for each dwelling. In some cases, researchers have had to resort to web scraping processes to obtain the large volumes of information that permit robust analyses [@gupta2022take; @arbia2020spatial; @Li2019; @lopez2015].  These  web  scraping  processes  can  include  missing  data, download errors, duplicate records, etc. Furthermore, the authors of this research do not generally share the data sets.

We are also witnessing a growing interest in open data in geography and data science [@arribasl2021editorial; @arribas2021]  using reproducible or replicable  research  [@paez2021open].  But  to  work  openly in  science, it  is  necessary to  have  free software and open data. While great efforts have been made to make free software available to researchers (e.g., R or Python), not much data is currently out in the open. In the particular case of the real estate market, to our knowledge, there are few open micro-data sets of housing markets available [@Song2021].

To overcome these limitations, this paper presents a sort description of an open micro-data set of geo-referenced dwelling listings. The data have been provided by the Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information about 189,923 dwellings located in Spain’s three largest cities. To date, this data product is the biggest open geo-referenced micro-data set of the housing market in Spain. Moreover, the data set has been supplied directly by Idealista, and therefore is clean and free of download errors. The listings have been enriched with official information from the Spanish cadastre along with other relevant geographical features, such as distance to urban points of interest. The data set is distributed as an R package, named ‘idealista18’, which can be accessed from the Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.

<!-- @cohen2008spatial (508 observations) -->
<!-- @lazrak2014market (19000 observations) -->

@@

The open data set ‘idealista18’ is an R package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia. For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the R package. The following subsections describe each object. A full description of the data is available in the help section of the package.

@@

The dwelling listing of each city includes a set of characteristics for each dwelling published on the idealista real estate website as an ad. The dwelling listing has been included in the ‘idealista18’ package as an sf object [@Pebesma]. The name of the sf object containing the dwelling listing includes the name of the city, followed by '_Sale' (e.g., Madrid_Sale) and includes a total of 42 variables. Each sf object includes the complete set of listings corresponding to the four quarters of the year 2018. Table \ref{tab:number-ads} shows the number of dwelling listing ads included in the data set for each city and quarter. The record counts for each city are: 94,815 listings for Madrid, 61,486 for Barcelona, and 33,622 for Valencia. Note that the same dwelling may be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter. The variable ASSETID, included in the sf objects, is the unique identifier of the dwelling.

```{r number-ads, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
H <- t(cbind(table(Barcelona_Sale$PERIOD),table(Madrid_Sale$PERIOD),table(Valencia_Sale$PERIOD)))
H <- cbind(H,rowSums(H))
text_tbl <- data.frame(
City = c("Barcelona","Madrid","Valencia"),
H,
check.names = FALSE)
colnames(text_tbl) <- c("City\\Quarter", "First", "Second ", "Thirdr", "Fourth","Total ads")
xtable::xtable(text_tbl, digits=0, caption = "Number of dwelling  listing ads for each city and quarter. \\label{tab:number-ads}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F) %>%
column_spec(1, width = "4em", italic = FALSE) %>%
column_spec(2, width = "3em") %>%
column_spec(3, width = "3em", latex_column_spec = "c") %>%
column_spec(4, width = "3em", latex_column_spec = "c") %>%
column_spec(5, width = "3em", latex_column_spec = "c") %>%
column_spec(6, width = "2em", latex_column_spec = "c")
```

Each record of the dwelling listing contains a set of indoor characteristics supplied by the advertisers on the Idealista website (e.g., price, surface area, number of bedrooms, basic features, etc.), including the exact location of the dwelling (see Section [Anonymizing the data set](#anonymizing)). Table \ref{tab:variables} lists the main indoor variables with a short description and the mean value of each variable. The dwelling listings were enriched with a number of additional attributes from the Spanish cadastre [@Catastro]. The cadastral information is described in Table \ref{tab:variables}, with the prefix CAD in the variable name. The cadastral features were assigned by applying the features of the nearest parcel to the coordinates. The year the dwelling was built (CONSTRUCTIONYEAR) given by the advertiser was revised since the year of construction is entered on the website by users, and it is therefore subject to errors and incomplete data (40\% missing data). To resolve this issue, an alternative variable (CADCONSTRUCTIONYEAR) was included, assigning the cadastral construction year from the nearest cadastral parcel whenever the value was outstanding (date was after publication date or year of construction was before 1500) or when the value supplied by the advertiser was missing.

Additionally, the distance of each dwelling to three urban points of interest was included in the sf object: distance to the city center, distance to the closest metro station, and distance to a major street (La Diagonal for Barcelona, La Castellana for Madrid, and Blasco Ibañez for Valencia). The last rows of Table \ref{tab:variables} show the mean values of these variables.

```{r table-variables, echo = FALSE, message = FALSE}
dummyvariables <- c(3,4,5,6,7,27,30,31,32,33,37,38,39)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("Asking price","Asking price per m^2 (euros)","Surface (m^2)",
                       "Number of bedrooms","Number of bathrooms",
                       "Construction year (advertiser)","Construction year (cadastre)",
                       "Max build floor","Dwelling count in the building",
                       "Cadastral quality. 0 Best-10 Worst",
                       "Distance to city center","Distance to subway station","Distance to major street"),
# Barcelona
Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables]), na.rm = TRUE),
# Madrid
Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables]), na.rm = TRUE),
# Valencia
"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables]), na.rm = TRUE),
check.names = FALSE)
text_tbl[dim(text_tbl)[1],1]<-"DISTANCE_TO_(MAINSTREET)"
xtable::xtable(text_tbl, caption = "List, sort description, and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities. See the help section in the \\textbf{idealista18} R package for details and formal definitions. Some variables have been excluded from this table to save space. Check the full list in the \\textbf{idealista18} package.\\label{tab:variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F, font_size=8) %>%
column_spec(1, width = "13em", italic = FALSE) %>%
column_spec(2, width = "14em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```

In addition to the variables listed in Table \ref{tab:variables}, the sf object includes a set of dummy variables with information about the basic characteristics of the dwelling. Table \ref{tab:Dummy-variables} shows the more relevant variables included in the sf object.

```{r table-dummy-variables, echo = FALSE, message = FALSE}
dummyvariables <- c(8,9,10,12,15,16,17,18,19,20,21,22,23,24,25,26,34,35,36)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("=1 if has terrace","=1 if has lift",
                       "=1 if has air conditioning","=1 if has parking",
                       "=1 if has north orientation","=1 if has south orientation",
                       "=1 if has east orientation","=1 if has west orientation",
                       "=1 if has boxroom","=1 if has wardrobe",
                       "=1 if has swimmingpool","=1 if has doorman",
                       "=1 if has garden","=1 if is duplex",
                       "=1 if is studio","=1 is on the top floor",
                       "=1 if is new contruction","=1 is second hand to be restored","=1 is second hand in good condition"),
# Barcelona
Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
# Madrid
Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables])),
# Valencia
"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables])),
check.names = FALSE)

xtable::xtable(text_tbl, caption = "List of dummy variables, sort description, and ratios of dwellings with the specific characteristics. See the help section in the \\textbf{idealista18} R package for details and formal definitions. Some dummy variables have been excluded from this table to save space\\label{tab:Dummy-variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F, font_size = 8) %>%
column_spec(1, width = "12em", italic = FALSE) %>%
column_spec(2, width = "14em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```

@@

The second block of data included in the ‘idealista18’ R package is the spatial features of the three cities divided into neighborhoods. There is an sf object for each city with the name of the city and the suffix '_Polygons'. Figure \ref{fig:all-polygons} shows the quantile maps of the number of dwellings in the listing for the different neighborhoods in the three cities. The neighborhoods are based on the official boundaries but slightly changed by Idealista\footnote{There are two criteria used to make this division. If an area is small enough and similar enough to another, the two areas are merged, and, if the official area is not homogeneous, it is divided into a series of new polygons.}. In practical terms, we can assume they are the same since the website combines areas when there are few ads for that area. In the case of Madrid, they combined four areas into two.

```{r, echo = FALSE,message=FALSE,fig.width=5, fig.cap="\\label{fig:all-polygons}Quantile maps of the number of dwellings in each neighborhood. Boundary for Barcelona (Left), Madrid (Center), and Valencia (Right)."}
library("ggplot2")
library("dplyr")
library("gridExtra")
# Barcelona
Barcelona_counts <- Barcelona_Polygons %>% 
  mutate(counts = lengths(st_intersects(., Barcelona_Sale)))
q <- quantile(Barcelona_counts$counts)
Barcelona_Polygons$Quantile<- as.factor((Barcelona_counts$counts > q[2]) + (Barcelona_counts$counts > q[3]) +(Barcelona_counts$counts >= q[4]) + 1)
plot4 <- ggplot(data = Barcelona_Polygons) +
  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
# Madrid
Madrid_counts <- Madrid_Polygons %>% 
  mutate(counts = lengths(st_intersects(., Madrid_Sale)))
q <- quantile(Madrid_counts$counts)
Madrid_Polygons$Quantile<- as.factor((Madrid_counts$counts > q[2]) + (Madrid_counts$counts > q[3]) +(Madrid_counts$counts >= q[4]) + 1)
plot5 <- ggplot(data = Madrid_Polygons) +
  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
# Valencia
Valencia_counts <- Valencia_Polygons %>% 
  mutate(counts = lengths(st_intersects(., Valencia_Sale)))
q <- quantile(Valencia_counts$counts)
Valencia_Polygons$Quantile<- as.factor((Valencia_counts$counts > q[2]) + (Valencia_counts$counts > q[3]) +(Valencia_counts$counts >= q[4]) + 1)
plot6 <- ggplot(data = Valencia_Polygons) +
  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
  ylim(39.405,39.52) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
grid.arrange(plot4, plot5, plot6, ncol = 3, nrow = 1)
```

There are a total of 69 neighborhoods in Barcelona, 135 in Madrid, and 73 in Valencia. The sf object includes a unique identifier (LOCATIONID) and the neighborhood name (LOCATIONNAME).

@@

The last block of data included in the data package is a set of Points of Interest in each city as an object of the class list. The name of the list includes the name of the city with the suffix '_POIS'. These lists include three elements: (i) the coordinates of the city center, the central business district; (ii) a set of coordinates that define the main street of each city; and (iii) the coordinates of the metro stations.

# Anonymizing the data set {#anonymizing}

To comply with Spanish regulations, two variables were slightly modified to provide anonymity. A masking process was applied to asking prices and location (coordinates).

In terms of the asking prices, the original values were obfuscated with the addition or subtraction of a random percentage of their original values, ranging from -2.5% to +2.5%. Since asking prices are usually multiples of 1,000, after the first price modification, the prices were aligned to multiples of 1,000.

@@

With respect to the location of the dwelling, a spatial masking process was implemented  to maintain the spatial properties of the original data set. The coordinates of each listing were displaced using a stochastic procedure. The listings were recorded using coordinates contained in maximum and minimum displacement circles, as shown in Figure \ref{fig:Anonymizing} (left). To preserve inclusion in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates remained in the original neighborhood of the listing.

Algorithm \ref{algo:coordinates-displacement} iteratively displaces the coordinates of each listing with a minimum distance and a maximum distance with the restriction that the new coordinates do not fall into a different neighborhood. This ensures that neighborhood attributes are preserved.

Figure \ref{fig:Anonymizing} (right) shows the histogram of the displacements in meters for all the listings in the city of Valencia. The average distance between the original and masked coordinates is 45 meters.

```{r, echo = FALSE, out.width=c("29%","37%"), out.height="20%",fig.cap="\\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Right) Coordinate displacement in meters (Valencia)",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("EPB_files/points-moved-image.png","EPB_files/coordinates-valencia.png"))
```

# Conclusion

This paper describes a data product of a geo-referenced micro-data set of Spain’s three largest cities. This is an excellent data product to help understand the complex mechanisms related to the housing market and housing prices. Researchers can apply hedonic models with spatial effects, identifying housing submarkets or machine learning techniques. The data product can also be used for educational proposes and teaching activities.

# Declaration of Competing Interest

The authors wish to thank Alessandro Galesi for his support in the paper revision and  Juan  Ramón  Selva  for  collecting  and  cleaning  the  spatial  data.  This  work  has been partially funded by the Spanish Ministry of Economy and Competitiveness Grant PID2019-107800GB-100.
